\documentclass[10.5pt,letterpaper]{article}

\usepackage[margin=0.72in]{geometry}
\usepackage{fontspec}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage{tabularx}
\usepackage{titlesec}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{ragged2e}

\IfFontExistsTF{Times New Roman}
  {\setmainfont{Times New Roman}}
  {\setmainfont{TeX Gyre Termes}}
\IfFontExistsTF{Apple Symbols}
  {\newfontfamily{\iconfont}{Apple Symbols}}
  {\newfontfamily{\iconfont}{TeX Gyre Termes}}
\newcommand{\setcviconfonts}[1]{%
  \newfontfamily{\brandiconfont}[Path=#1]{fa-brands-400.ttf}%
  \newfontfamily{\solidiconfont}[Path=#1]{fa-solid-900.ttf}%
  \newfontfamily{\academiciconfont}[Path=#1]{academicons.ttf}%
}
\IfFileExists{../assets/fonts/fa-brands-400.ttf}
  {\setcviconfonts{../assets/fonts/}}
  {\IfFileExists{fonts/fa-brands-400.ttf}
    {\setcviconfonts{fonts/}}
    {\newfontfamily{\brandiconfont}{Arial Unicode MS}
     \newfontfamily{\solidiconfont}{Arial Unicode MS}
     \newfontfamily{\academiciconfont}{Arial Unicode MS}}}

\definecolor{ink}{HTML}{1F2933}
\definecolor{muted}{HTML}{52606D}
\definecolor{rule}{HTML}{CBD2D9}
\definecolor{accent}{HTML}{1F4E79}

\hypersetup{
  colorlinks=true,
  urlcolor=accent,
  linkcolor=accent
}

\pagestyle{empty}
\setlength{\parindent}{0pt}
\setlength{\parskip}{2pt}
\setlist[itemize]{leftmargin=1.15em, topsep=2pt, itemsep=2pt, parsep=0pt}
\setlist[enumerate]{leftmargin=1.35em, topsep=2pt, itemsep=2pt, parsep=0pt}

\titleformat{\section}
  {\large\scshape\color{accent}}
  {}
  {0pt}
  {}
  [\vspace{-3pt}\color{rule}\titlerule]
\titlespacing{\section}{0pt}{9pt}{5pt}

\newcommand{\cvsubtle}[1]{{\color{muted}#1}}
\newcommand{\role}[1]{\textit{#1}}
\newcommand{\cvstrong}[1]{\textbf{#1}}
\newcommand{\cvicon}[1]{\raisebox{-0.05ex}{\scriptsize\color{accent}#1}}
\newcommand{\cvmutedicon}[1]{\raisebox{-0.05ex}{\scriptsize\color{muted}#1}}
\newcommand{\faGithub}{{\brandiconfont\symbol{"F09B}}}
\newcommand{\faDiscord}{{\brandiconfont\symbol{"F392}}}
\newcommand{\faEnvelope}{{\solidiconfont\symbol{"F0E0}}}
\newcommand{\faPhone}{{\solidiconfont\symbol{"F095}}}
\newcommand{\faMapMarker}{{\solidiconfont\symbol{"F3C5}}}
\newcommand{\faUsers}{{\solidiconfont\symbol{"F0C0}}}
\newcommand{\faMusic}{{\solidiconfont\symbol{"F001}}}
\newcommand{\faLink}{{\solidiconfont\symbol{"F0C1}}}
\newcommand{\faFilePdf}{{\solidiconfont\symbol{"F1C1}}}
\newcommand{\aiGoogleScholar}{{\academiciconfont\symbol{"E828}}}
\newcommand{\aiArxiv}{{\academiciconfont\symbol{"E808}}}
\newcommand{\linkglyph}{\,{\cvicon{\faLink}}}
\newcommand{\link}[2]{\href{#1}{#2\linkglyph}}
\newcommand{\contactlink}[3]{\href{#1}{\cvicon{#2}\hspace{0.22em}#3}}
\newcommand{\mailcontactlink}[2]{\contactlink{#1}{\faEnvelope}{#2}}
\newcommand{\papericonlink}[2]{\href{#1}{\cvicon{#2}}}
\newcommand{\paperlinks}[1]{\hspace{0.25em}{#1}}
\newcommand{\daterange}[1]{{\color{muted}#1}}
\newcommand{\entry}[4]{%
  \begin{tabularx}{\linewidth}{@{}Xr@{}}
    \textbf{#1} & \daterange{#2} \\
    \textit{#3} & \cvsubtle{#4}
  \end{tabularx}
}
\newcommand{\projectentry}[3]{%
  \begin{tabularx}{\linewidth}{@{}Xr@{}}
    \textbf{#1} & \daterange{#2}
  \end{tabularx}
  #3
}

\begin{document}

{\huge \textbf{Ruibin Yuan}}\\[-1pt]
{\large PhD Student in Artificial Intelligence, HKUST \, | \, AI Researcher, Developer, and Musician}

\vspace{4pt}
\begin{tabularx}{\linewidth}{@{}Xr@{}}
\mailcontactlink{mailto:ruibiny@alumni.cmu.edu}{ruibiny@alumni.cmu.edu} \, | \,
\mailcontactlink{mailto:a43992899@gmail.com}{a43992899@gmail.com}
&
\cvmutedicon{\faMapMarker}\hspace{0.22em}Hong Kong
\\
\contactlink{https://github.com/a43992899}{\faGithub}{github.com/a43992899} \, | \,
\contactlink{https://scholar.google.com/citations?user=Qd_hX1cAAAAJ}{\aiGoogleScholar}{Google Scholar} \, | \,
\contactlink{https://huggingface.co/m-a-p}{\faUsers}{MAP Research Community}
&
\cvmutedicon{\faPhone}\hspace{0.22em}+852 46830974 \, | \, +86 18033231161
\end{tabularx}

\vspace{3pt}
\color{rule}\hrule
\color{ink}
\vspace{6pt}

\textbf{Research focus.} Foundation models for music generation and understanding; multimodal LLMs; large-scale data, training infrastructure, and open research communities for creative AI.

\section{Education}

\entry{Hong Kong University of Science and Technology (HKUST)}{09/2023--Present}
{PhD in Artificial Intelligence}{Hong Kong}
Supervisors: Jie Fu, Wei Xue, Yike Guo

\entry{Carnegie Mellon University (CMU)}{09/2021--05/2023}
{M.S. in Music and Technology, Computer Science Emphasis}{Pittsburgh, PA}
Supervisors: Roger B. Dannenberg, Gus Xia, Riccardo Schulz

\entry{Beijing University of Posts and Telecommunications (BUPT)}{09/2016--06/2020}
{B.S. in Electronic Science and Technology}{Beijing, China}

\section{Academic Experience}

\projectentry{Qwen Music}{04/2025--Present}{
\textit{Team Lead.} Lead foundation model research for music generation and contribute music understanding capabilities in the Qwen-Omni series.
}

\projectentry{\link{https://huggingface.co/m-a-p}{Multimodal Art Projection Research Community (MAP)}}{09/2022--Present}{
\textit{Cofounder.} Build open large language, music, and multimodal models, with a focus on open-source training, data collection, and creative AI. Lead MAP's multimodal and AI music research direction across projects including COIG, MERT, ChatMusician, CMMMU, and YuE.
}

\projectentry{\link{https://github.com/multimodal-art-projection/YuE}{YuE (OpenSuno)}}{03/2024--03/2025}{
\textit{Team Lead.} Built an open full-song generation foundation model for lyrics-to-song and long-form music generation, attracting broad open-source adoption.
}

\projectentry{\link{https://huggingface.co/collections/m-a-p/neo-models-66395a5c9662bb58d5d70f04}{NEO LLM}}{09/2023--05/2024}{
\textit{Core Member.} Helped train a fully open LLM from scratch, contributing to infrastructure, PB-scale data collection and cleaning, training, evaluation, experiment logging, and tooling. Operated large GPU clusters and monitored hardware/software failures during training.
}

\projectentry{\link{https://huggingface.co/collections/m-a-p/chatmusician-65de07b3b87b189c2a588329}{ChatMusician}}{2023--2024}{
\textit{Team Lead.} Trained the first symbolic music LLM; collected a large-scale symbolic music and language dataset and an advanced music understanding benchmark. The project received public attention from Yann LeCun and AK.
}

\projectentry{\link{https://huggingface.co/datasets/MMMU/MMMU}{MMMU \& CMMMU}}{2023--2024}{
Contributed to multimodal benchmark development extending MMLU/CMMLU-style expert reasoning to multimodal settings. MMMU was recommended by Jeff Dean.
}

\projectentry{\link{https://huggingface.co/collections/m-a-p/mert-65a2cb498448f47df2d9e927}{MERT}}{2022--2023}{
Built a large-scale self-supervised BERT-style representation model for music audio.
}

\projectentry{\link{https://marble-bm.shef.ac.uk}{MARBLE}}{2022--2023}{
Built the first unified benchmark for music understanding.
}

\projectentry{\link{https://huggingface.co/collections/m-a-p/coig-65adf31cd6b4c93ba6894def}{COIG}}{2022}{
\textit{Project Lead, COIG-PC.} Led the Chinese FLAN-style instruction tuning corpus effort, coordinating over one hundred contributors and converting thousands of pre-LLM-era Chinese NLP datasets into instruction format.
}

\section{Selected Publications and Manuscripts}

\textit{Asterisks indicate equal contribution. Icons link to papers, code, or project pages.}

{\small
\begin{enumerate}[itemsep=1pt]
\item \textbf{[arXiv 2025]} YuE: Scaling Open Foundation Models for Long-Form Music Generation \paperlinks{\papericonlink{https://arxiv.org/abs/2503.08638}{\aiArxiv}\,\papericonlink{https://github.com/multimodal-art-projection/YuE}{\faGithub}}. \cvstrong{Ruibin Yuan}, Hanfeng Lin, Shuyue Guo, Ge Zhang, Jiahao Pan, et al.

\item \textbf{[CVPR 2024, highly cited]} MMMU: A Massive Multi-discipline Multimodal Understanding and Reasoning Benchmark for Expert AGI \paperlinks{\papericonlink{https://doi.org/10.1109/cvpr52733.2024.00913}{\faFilePdf}}. Xiang Yue*, Yuansheng Ni*, Kai Zhang*, Tianyu Zheng*, et al., \cvstrong{Ruibin Yuan}, et al.

\item \textbf{[ICLR 2024]} MERT: Acoustic Music Understanding Model with Large-Scale Self-supervised Training \paperlinks{\papericonlink{https://arxiv.org/abs/2306.00107}{\aiArxiv}}. Yizhi Li*, \cvstrong{Ruibin Yuan}*, Ge Zhang*, Yinghao Ma*, et al.

\item \textbf{[NeurIPS 2023]} MARBLE: Music Audio Representation Benchmark for Universal Evaluation \paperlinks{\papericonlink{https://arxiv.org/abs/2306.10548}{\aiArxiv}\,\papericonlink{https://marble-bm.shef.ac.uk}{\faMusic}}. \cvstrong{Ruibin Yuan}*, Yinghao Ma*, Yizhi Li*, Ge Zhang*, et al.

\item \textbf{[ACL 2024 Main, highly cited]} AnyGPT: Unified Multimodal LLM with Discrete Sequence Modeling \paperlinks{\papericonlink{https://doi.org/10.18653/v1/2024.acl-long.521}{\faFilePdf}}. Jun Zhan, Junqi Dai, Jiasheng Ye, Yunhua Zhou, et al., \cvstrong{Ruibin Yuan}, et al.

\item \textbf{[COLM 2024, highly cited]} RQ-RAG: Learning to Refine Queries for Retrieval Augmented Generation \paperlinks{\papericonlink{https://openreview.net/forum?id=tzE7VqsaJ4}{\faFilePdf}\,\papericonlink{https://github.com/chanchimin/RQ-RAG}{\faGithub}}. Chi-Min Chan, Chunpu Xu, \cvstrong{Ruibin Yuan}, Hongyin Luo, Wei Xue, Yike Guo, Jie Fu.

\item \textbf{[ICLR 2025]} MuPT: A Generative Symbolic Music Pretrained Transformer \paperlinks{\papericonlink{https://proceedings.iclr.cc/paper_files/paper/2025/hash/73f6f8897896f7bda86ea7d1ebc1dc4f-Abstract-Conference.html}{\faFilePdf}}. Xingwei Qu, Yuelin Bai, Yinghao Ma, Ziya Zhou, Ka Man Lo, Jiaheng Liu, \cvstrong{Ruibin Yuan}, et al.

\item \textbf{[ACL 2025 Findings]} CLaMP 3: Universal Music Information Retrieval Across Unaligned Modalities and Unseen Languages \paperlinks{\papericonlink{https://aclanthology.org/2025.findings-acl.133/}{\faFilePdf}\,\papericonlink{https://github.com/sanderwood/clamp3}{\faGithub}}. Shangda Wu, Zhancheng Guo, \cvstrong{Ruibin Yuan}, Junyan Jiang, Seungheon Doh, Gus Xia, Juhan Nam, Xiaobing Li, Feng Yu, Maosong Sun.

\item \textbf{[NAACL 2025 Findings]} CLaMP 2: Multimodal Music Information Retrieval Across 101 Languages Using Large Language Models \paperlinks{\papericonlink{https://aclanthology.org/2025.findings-naacl.27/}{\faFilePdf}}. Shangda Wu, Yashan Wang, \cvstrong{Ruibin Yuan}, Zhancheng Guo, Xu Tan, Ge Zhang, et al.

\item \textbf{[ACL 2024 Findings]} ChatMusician: Understanding and Generating Music Intrinsically with LLM \paperlinks{\papericonlink{https://arxiv.org/abs/2402.16153}{\aiArxiv}}. \cvstrong{Ruibin Yuan}, Hanfeng Lin, Yi Wang, Zeyue Tian, Shangda Wu, Tianhao Shen, Ge Zhang, et al.

\item \textbf{[arXiv 2024]} CMMMU: A Chinese Massive Multi-discipline Multimodal Understanding Benchmark \paperlinks{\papericonlink{https://arxiv.org/abs/2401.11944}{\aiArxiv}}. Ge Zhang, Xinrun Du, Bei Chen, Yiming Liang, Tongxu Luo, Tianyu Zheng, et al., \cvstrong{Ruibin Yuan}, et al.

\item \textbf{[NAACL 2025 Findings]} COIG-CQIA: Quality is All You Need for Chinese Instruction Fine-tuning \paperlinks{\papericonlink{https://aclanthology.org/2025.findings-naacl.457/}{\faFilePdf}\,\papericonlink{https://huggingface.co/datasets/m-a-p/COIG-CQIA}{\faUsers}}. Yuelin Bai, Xeron Du, Yiming Liang, Leo Jin, Junting Zhou, et al., \cvstrong{Ruibin Yuan}, et al.

\item \textbf{[arXiv 2024]} OmniBench: Towards The Future of Universal Omni-Language Models \paperlinks{\papericonlink{https://arxiv.org/abs/2409.15272}{\aiArxiv}}. Yizhi Li, Ge Zhang, Yinghao Ma, \cvstrong{Ruibin Yuan}, Kang Zhu, Hangyu Guo, et al.

\item \textbf{[ISMIR 2023]} LyricWhiz: Robust Multilingual Zero-shot Lyrics Transcription by Whispering to ChatGPT \paperlinks{\papericonlink{https://arxiv.org/abs/2306.17103}{\aiArxiv}}. Le Zhuo, \cvstrong{Ruibin Yuan}, Jiahao Pan, Yinghao Ma, Yizhi Li, Ge Zhang, et al.

\item \textbf{[ISMIR 2023]} On the Effectiveness of Speech Self-supervised Learning for Music \paperlinks{\papericonlink{https://arxiv.org/abs/2307.05161}{\aiArxiv}}. Yinghao Ma, \cvstrong{Ruibin Yuan}, Yizhi Li, Ge Zhang, Xingran Chen, Hanzhi Yin, et al.

\item \textbf{[Interspeech 2022]} DeID-VC: Speaker De-identification via Zero-shot Pseudo Voice Conversion \paperlinks{\papericonlink{https://doi.org/10.21437/interspeech.2022-11036}{\faFilePdf}}. \cvstrong{Ruibin Yuan}, Yuxuan Wu, Jacob Li, Jaxter Kim.
\end{enumerate}
}

\section{Industry Experience}

\entry{\link{https://chat.qwen.ai/}{Qwen}}{04/2025--Present}
{Research Intern, Qwen Music; Supervisor: \link{https://scholar.google.com/citations?user=COYDNmYAAAAJ}{Jin Xu}}{Remote}

\entry{\link{https://kimi.ai}{Moonshot.ai}}{09/2024--04/2025}
{Research Intern; Supervisor: \link{https://scholar.google.com/citations?user=tob-U1oAAAAJ}{Xu Tan}}{Remote}

\entry{\link{https://huggingface.co/BAAI}{BAAI}}{2022--2023}
{Part-time Research Intern; Supervisor: \link{https://scholar.google.com/citations?user=66osleIAAAAJ}{Jie Fu}}{Remote}

\entry{\link{https://stardust.ai}{Stardust.ai}}{05/2020--2021}
{Part-time ML Consultant; Ex-engineer}{Beijing, China}

\entry{\link{https://music.163.com}{NetEase, Inc.}}{12/2019--04/2020}
{Intern Machine Learning Engineer}{Shanghai, China}

\entry{Vlahakis Recording Studio at CMU}{2023}
{Student Audio Engineer}{Pittsburgh, PA}

\section{Music Experience}

\begin{tabularx}{\linewidth}{@{}Xr@{}}
Lead Guitarist and Cofounder, BUPT Slow Blues Band \href{https://music.163.com/\#/album?id=87161044}{[Album]} & \daterange{2016--2021} \\
President, BUPT Guitar Club & \daterange{2017--2018}
\end{tabularx}

\end{document}
